---
title: "Recitation"
author: "Mohammed Ali"
date: "July 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Read in the data
```{r}
NBA <- read.csv("data/NBA_train.csv")
glimpse(NBA)
```

# How many wins to make the playoffs?
```{r}
table(NBA$W, NBA$Playoffs)
```

# Compute Points Difference
```{r}
NBA$PTSdiff <- NBA$PTS - NBA$oppPTS
```

# Check for linear relationship
```{r}
plot(NBA$PTSdiff, NBA$W)
```

# Linear regression model for wins
```{r}
WinsReg <- lm(W ~ PTSdiff, data=NBA)
summary(WinsReg)
```

# Linear regression model for points scored
```{r}
PointsReg <- lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + DRB + TOV + STL + BLK, data=NBA)
summary(PointsReg)
```

# Sum of Squared Errors
```{r}
PointsReg$residuals
SSE <- sum(PointsReg$residuals^2)
SSE
```

# Root mean squared error
```{r}
RMSE <- sqrt(SSE/nrow(NBA))
RMSE
```

# Average number of points in a season
```{r}
mean(NBA$PTS)
```


# Remove insignifcant variables
```{r}
summary(PointsReg)

PointsReg2 <- lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + DRB + STL + BLK, data=NBA)
summary(PointsReg2)

PointsReg3 <- lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + STL + BLK, data=NBA)
summary(PointsReg3)

PointsReg4 <- lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + STL, data=NBA)
summary(PointsReg4)
```

# Compute SSE and RMSE for new model
```{r}
SSE_4 <- sum(PointsReg4$residuals^2)
RMSE_4 <- sqrt(SSE_4/nrow(NBA))
SSE_4
RMSE_4
```


# Read in test set
```{r}
NBA_test <- read.csv("data/NBA_test.csv")
```

# Make predictions on test set

```{r}
PointsPredictions = predict(PointsReg4, newdata=NBA_test)
```

# Compute out-of-sample R^2
```{r}
SSE <- sum((PointsPredictions - NBA_test$PTS)^2)
SST <- sum((mean(NBA$PTS) - NBA_test$PTS)^2)
R2 <- 1 - SSE/SST
R2
```

# Compute the RMSE
```{r}
RMSE <- sqrt(SSE/nrow(NBA_test))
RMSE
```
