---
title: "Boston Housing Data"
author: "Mohammed Ali"
date: "July 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caTools)
library(rpart)
library(rpart.plot)
library(ROCR)
library(randomForest)
library(caret)
library(e1071)
```

# EDA

## Read in data
```{r}
boston <- read.csv("data/boston.csv")
glimpse(boston)
```
## Plot observations
```{r}
ggplot(boston, aes(x = LON, y = LAT)) +
  geom_point()
```

## Tracts alongside the Charles River
```{r}
chas <- boston[boston$CHAS==1,c("LON", "LAT")]

ggplot() +
  geom_point(data = boston, aes(x = LON, y = LAT)) +
  geom_point(data = chas, aes(x = LON, y = LAT), col="blue",shape = 0) 
```

## Plot MIT

```{r}
mit <- boston[boston$TRACT==3531,c("LON", "LAT")]

ggplot() +
  geom_point(data = boston, aes(x = LON, y = LAT)) +
  geom_point(data = mit, aes(x = LON, y = LAT), col="red", size = 3, shape = 2) 
```


## Plot polution

```{r}
poll <- boston[boston$NOX>=0.55, c("LON", "LAT")]

ggplot() +
  geom_point(data = boston, aes(x = LON, y = LAT)) +
  geom_point(data = poll, aes(x = LON, y = LAT), col="red", size = 2, shape = 2)  
```

## Plot prices

```{r}
prices <- boston[boston$MEDV>=21.2,  c("LON", "LAT")]

ggplot() +
  geom_point(data = boston, aes(x = LON, y = LAT)) +
  geom_point(data = prices, aes(x = LON, y = LAT), col="green", size = 2, shape = 3) 
```


# Linear Regression using LAT and LON
```{r}
plot(boston$LAT, boston$MEDV)
plot(boston$LON, boston$MEDV)
latlonlm = lm(MEDV ~ LAT + LON, data=boston)
summary(latlonlm)
```

## Visualize regression output
```{r}
plot(boston$LON, boston$LAT)
points(boston$LON[boston$MEDV>=21.2], boston$LAT[boston$MEDV>=21.2], col="red", pch=20)

latlonlm$fitted.values
points(boston$LON[latlonlm$fitted.values >= 21.2], boston$LAT[latlonlm$fitted.values >= 21.2], col="blue", pch="$")
```


# Regression Trees

## CART model
```{r}
latlontree = rpart(MEDV ~ LAT + LON, data=boston)
prp(latlontree)
```

## Visualize output
```{r}
plot(boston$LON, boston$LAT)
points(boston$LON[boston$MEDV>=21.2], boston$LAT[boston$MEDV>=21.2], col="red", pch=20)

fittedvalues = predict(latlontree)
points(boston$LON[fittedvalues>21.2], boston$LAT[fittedvalues>=21.2], col="blue", pch="$")
```



## Simplify tree by increasing minbucket
```{r}
latlontree = rpart(MEDV ~ LAT + LON, data=boston, minbucket=50)
plot(latlontree)
text(latlontree)
```


## Visualize Output
```{r}
plot(boston$LON,boston$LAT)
abline(v=-71.07)
abline(h=42.21)
abline(h=42.17)
points(boston$LON[boston$MEDV>=21.2], boston$LAT[boston$MEDV>=21.2], col="red", pch=20)
```


# Let's use all the variables

## Split the data
```{r}
set.seed(123)
split = sample.split(boston$MEDV, SplitRatio = 0.7)
train = subset(boston, split==TRUE)
test = subset(boston, split==FALSE)
```

## Create linear regression
```{r}
linreg = lm(MEDV ~ LAT + LON + CRIM + ZN + INDUS + CHAS + NOX + RM + AGE + DIS + RAD + TAX + PTRATIO, data=train)
summary(linreg)
```

## Make predictions
```{r}
linreg.pred = predict(linreg, newdata=test)
linreg.sse = sum((linreg.pred - test$MEDV)^2)
linreg.sse
```

## Create a CART model
```{r}
tree = rpart(MEDV ~ LAT + LON + CRIM + ZN + INDUS + CHAS + NOX + RM + AGE + DIS + RAD + TAX + PTRATIO, data=train)
prp(tree)
```

## Make predictions
```{r}
tree.pred = predict(tree, newdata=test)
tree.sse = sum((tree.pred - test$MEDV)^2)
tree.sse
```


# Cross Validation

## Number of folds
```{r}
tr.control = trainControl(method = "cv", number = 10)
```

## cp values
```{r}
cp.grid = expand.grid( .cp = (0:10)*0.001)
```

### What did we just do?
`1*0.001` 
`10*0.001`
`0:10`
`0:10 * 0.001`


## Cross-validation
```{r}
tr = train(MEDV ~ LAT + LON + CRIM + ZN + INDUS + CHAS + NOX + RM + AGE + DIS + RAD + TAX + PTRATIO, data = train, method = "rpart", trControl = tr.control, tuneGrid = cp.grid)
```

## Extract tree
```{r}
best.tree = tr$finalModel
prp(best.tree)
```

## Make predictions
```{r}
best.tree.pred = predict(best.tree, newdata=test)
best.tree.sse = sum((best.tree.pred - test$MEDV)^2)
best.tree.sse
```
